---
title: "Rethinking_week04_2019-04-25.Rmd"
author: "Stacey Harmer"
date: "4/26/2019"
output: 
  html_document: 
    keep_md: yes
---

## Video: Lecture 3

Quadratic approximation - can use small amount of informaiton to define entire curve.  

quap uses MAP
```{r}
dunif(51, 0,50)
dunif(51, 0, 51)
```
log-normal distrbution (  rlnorm )
resuies all sampled values be positive.  

```{r}

x <- rlnorm(1000, 0, 1)
dens(x)
dens(log(x))
```


## Reading: Chapter 4 through section 4.4.2  / page 98

Problems: 4E1-4E5; 4M1, 4M2, 4M3

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

By “linear regression,”
we will mean a family of simple statistical golems that attempt to learn about the mean and
variance of some measurement, using an additive combination of other measurements.

## 4.1  normal distibutions

```{r}
?runif

pos <- replicate(1000, sum(runif(16, -1, 1)))
hist(pos)
plot(density(pos))

pos <- replicate(1000, sum(runif(1e4, -1, 1))^2)
hist(pos)
plot(density(pos))

# that is weird, but makes sense because centered near zero.


pos <- replicate(1000, sum(runif(1e4, -1, 1)^2))
hist(pos)
plot(density(pos))
# back to normal
```


multiplicative

```{r}
library(rethinking)
growth <- replicate( 10000 , prod( 1 + runif(12,0,0.1) ) )
dens( growth , norm.comp=TRUE )

```

note that:
Multiplying small numbers is approximately the same as addition. For
example, if there are two loci with alleles increasing growth by 10% each, the product is:


```{r}

big <- replicate( 10000 , prod( 1 + runif(12,0,0.5) ) )  # effects up to 0.5
dens(big)
small <- replicate( 10000 , prod( 1 + runif(12,0,0.01) ) )  # effects up to 0.01;  numbers close to 1!
dens(small)

```

log multilplication  - copes with big deviations

```{r}
log.big <- replicate( 10000 , log(prod(1 + runif(12,0,0.5))) )
dens(log.big)
# can I transform earlier result?
dens(log(big))  #Yep

```
Hey, note that 
And repeatedly adding finite fluctuations
results in a distribution of sums that have shed all information about the underlying process,
aside from mean and spread

Gaussian distribution has quadratic shape.

```{r}
curve( exp( -x^2 ) , from=-3 , to=3 )
curve(  -x^2  , from=-3 , to=3 )
```
keep in mind Gaussian is continuous, a probability density funciton denoted with p or even with f.
while binomial requires integer (and is a probability mass function, Pr)

he defines tau on page 76.  

## 4.2  Describing models.

data = what we can observe
parameters = unobservable things like rates and averages. 
variables = defined either in terms of other varialbles or in terms of prabilbitly distriution.  

generally, ~ means 'is distributed' 

```{r}
p_grid <- seq(from=0,to=1,length.out=100)
dunif(p_grid,0,1)  #I don't understand this part of section 4.6 (page 78)
?dunif

```

## 4.3  Gaussian model of height  (p 78)


```{r}
data(Howell1)
d <- Howell1
str(d)
```

```{r}
precis(d)
```

stopped page 80.


## homework questions
Problems: 4E1-4E5; 4M1, 4M2, 4M3
p 121

### 4E1.


