---
title: "Rethinking_week08_2019-06-04"
author: "Stacey Harmer"
date: "6/4/2019"
output: 
  html_document: 
    keep_md: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Reading: Finish chapter 6

# Video: First half of Ulysses' Compass <slides> <video> (through 36:30)

## Causal inference
Confounds: 1. fork, 2. pipe, (open until you condition on Z)
3. collider (closed until you condition on Z)  light switch example
4. descendent (conditioning on descendent is like condititionig on its parent)

"information can walk against the arrows" in a DAG.  causation can only flow with arrows
two arrows into a factor 'close' it, nullifying the collider. 

minute 19 - waffle house divorce DAG

interventions: can change many variables!  that commplicates studies.

stopped minute 36

# chapter 6

pairs plot - if 2 varaible havs correlation > 0.9, beware!

## 6.4   Confounding

code 6.30
```{r}
library(dagitty) 
library(rethinking)
dag_6.1 <- dagitty( "dag {
X -> Y <- C
X <- U -> B
U <- A -> C
U -> B <- C
}")
# plot(dag_6.1)  # need coordinates before plottin
adjustmentSets( dag_6.1 , exposure="X" , outcome="Y" )
```

code 6.31
```{r}
dag_6.2 <- dagitty( "dag {
S -> A -> D
S -> M -> D
S -> W -> D
A -> M
}")
adjustmentSets( dag_6.2 , exposure="W" , outcome="D" )
```
 code 6.32
```{r}
impliedConditionalIndependencies( dag_6.2 )
```
 Read the first as “median age of marriage should be independent of (_||_) Waffle Houses,
conditioning on (|) a State being in the south.”
 
# Problems
 
##6H1.  Use the Waffle House data, data(WaffleDivorce), 
to find the total causal influence of number of Waffle Houses on divorce rate. Justify your model or models with a causal graph.
 
OK, first make my causal graph.  Reproduce what is shown on p 184

```{r}
dag_6h1 <- dagitty( "dag {
S -> A -> D
S -> M -> D
S -> W -> D
A -> M
}")

coordinates( dag_6h1 ) <- list( x=c(S=0,A = 0, M = 1, W = 2, D = 2) ,
y=c(S = 0, W = 0, A = 2, M = 1, D = 2) )
plot( dag_6h1 )


```
I want to examine effects of waffle house on divorce.

According to chapter, that means I just condition on S first.  That is, I include S in my model, as well as M and D.

So load the data, then make my model
```{r}
data("WaffleDivorce")
waffle <- WaffleDivorce

```
standardize variables

```{r}

waffle$D <- scale(waffle$Divorce)
waffle$A <- scale(waffle$MedianAgeMarriage)
waffle$M <- scale(waffle$Marriage)
waffle$W <- scale(waffle$WaffleHouses)

#and to use notation he favors, make South an index variable
# 1 means not South, 2 means South

waffle$S <- ifelse( waffle$South==1 , 2 , 1 )

```

now I think I'm ready to make a model, coditioning on S

see pages 151/153 for use of indicator varaible

modified from page 130

Do I need an alpha at all?  Or can I just use S as the alpha?  p 149
I htink so - 2 alphas means I am assuming two different mean divorce rates, south and not south
 
divorce ~ normal(mu, sigma) 
mu = alpha[south] + beta.waffle(W)   

```{r}
m6h1 <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a[S] + beta.waffle*W , 
    a[S]  ~ dnorm( 0, 0.5),
    beta.waffle ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h1, depth = 2)

plot(precis(m6h1, depth = 2))
```
Remember 1 means not South, 2 means South
a trend for mean to be negative for not south, positive for south.  But crosses zero.

Beta.waffle mean is positive: 0.08.  slight positive effect on divorce?  But interval is from -0.17 to 0.34, so no confidence in its effect.


I'll remake my DAG, removing connection between W and D
```{r}
dag_6h1b <- dagitty( "dag {
S -> A -> D
S -> M -> D
S -> W 
A -> M
}")

coordinates( dag_6h1b ) <- list( x=c(S=0,A = 0, M = 1, W = 2, D = 2) ,
y=c(S = 0, W = 0, A = 2, M = 1, D = 2) )
plot( dag_6h1b )
```

## problem 6H2.  Build a series of models to test the implied conditional independencies 
of the causal graph you used in the previous problem. If any of the tests fail, how do you think the graph needs to be amended? Does the graph need more or fewer arrows? Feel free to nominate variables that aren’t in the data.

First, look at association between S and W
```{r}
m6h2.SW <- quap(
  alist(
    W ~ dnorm(mu, sigma),
    mu <- a[S]  , 
    a[S]  ~ dnorm( 0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h2.SW, depth = 2)

plot(precis(m6h2.SW, depth = 2))
```
That is clear.  strong positive association between S and Waffle house

Next, look at association between S and M
```{r}
m6h2.SM <- quap(
  alist(
    M ~ dnorm(mu, sigma),
    mu <- a[S]  , 
    a[S]  ~ dnorm( 0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h2.SM, depth = 2)

plot(precis(m6h2.SM, depth = 2))
```
Not sure we can draw that line between S and M.  I"ll remove from DAG

OK, what about S and A?
```{r}
m6h2.SA <- quap(
  alist(
    A ~ dnorm(mu, sigma),
    mu <- a[S]  , 
    a[S]  ~ dnorm( 0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h2.SA, depth = 2)

plot(precis(m6h2.SA, depth = 2))
``` 
This is also not entirely clear.  but bulk of associations are as suggested. Leave in the DAG

What about A to M?

```{r}
m6h2.AM <- quap(
  alist(
    M ~ dnorm(mu, sigma),
    mu <- beta.A*A  , 
    beta.A ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)
?WaffleDivorce
precis(m6h2.AM, depth = 2)

plot(precis(m6h2.AM, depth = 2))
``` 
beta.A is strongly negative; the greater the age,the lower the M?  Yep. as expected, really.

```{r}
plot(waffle$M, waffle$A)
```
What about A to D?

```{r}
m6h2.AD <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- beta.A*A  , 
    beta.A ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h2.AD, depth = 2)

plot(precis(m6h2.AD, depth = 2))
``` 
As above.

Shoudl I include a direct S to D??  If so, I need to control for A

```{r}
m6h2.SDA <- quap(
  alist(
    D ~ dnorm(mu, sigma),
    mu <- a[S] + beta.age*A , 
    a[S]  ~ dnorm( 0, 0.5),
    beta.age ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = waffle)

precis(m6h2.SDA, depth = 2)

plot(precis(m6h2.SDA, depth = 2))

# seems like South has posiitve effect on divorce, although it is a bit close,
# even after controlling for age
```

OK, I think I need to modify my DAG to remove S -> M
I will keep the S to A, though.  And I'll add S -> D

```{r}
dag_6h2.end <- dagitty( "dag {
S -> A -> D
S -> D
S -> W 
A -> M
}")

coordinates( dag_6h2.end ) <- list( x=c(S=0,A = 0, M = 0, W = 1, D = 1) ,
y=c(S = 0, W = 0, A = .5, M = 1, D = 1) )
plot( dag_6h2.end )
```

What if I considered the W to D connection using above graph?
I'd still just control on S.


And the three problems at https://github.com/rmcelreath/statrethinking_winter2019/blob/master/homework/week03.pdf

# Github homework - fox data

```{r}
library(rethinking)
data(foxes)
fox <- foxes
```
Want to model effect of area on weight.  
there is a pipe: area --> food --> weight
and there is a collider: food --> groupsize --> weight

## fox 1

I think he is asking us to simply model weight by area, ignoring other factors

First, scale data 

```{r}
fox$F <- scale(fox$avgfood)
fox$G <- scale(fox$groupsize)
fox$A <- scale(fox$area)
fox$W <- scale(fox$weight)
```
plot just to get sense
```{r}
plot(fox$A, fox$W)  # not a very good correpondence
pairs(fox) # food and group size are closely related, as are area and groupsize. 
# area and food also.  weight not clearly correlated to any one factor

```

Next, simple model with no additional factors.  Weight as influenced just by area.
```{r}
fox1 <- quap(
  alist(
    W ~ dnorm(mu, sigma),
    mu <- beta.area*A , 
    beta.area ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = fox)

precis(fox1)

plot(precis(fox1))
```
as expected from pairs plot, area doesn't appear explanatory for weight

Oops, I need to examine my priors.  see p 126 in book

```{r}
set.seed(10)
prior <- extract.prior( fox1 )
mu <- link( fox1 , post=prior , data=list( A=c(-2,2) ) )
plot( NULL , xlim=c(-2,2) , ylim=c(-2,2) )
for ( i in 1:50 ) lines( c(-2,2) , mu[i,] , col=col.alpha("black",0.4) )
```
I think this is OK; my input range is from -2.2 to 2
```{r}
summary(fox)
```
## fox 2
Now add food to the model. I think that i DO NOT want to add groupsize to the model as adding a collider is problematic.

model weight as effect of area and food availabile. 
```{r}
fox2 <- quap(
  alist(
    W ~ dnorm(mu, sigma),
    mu <- beta.area*A + beta.food*F, 
    beta.area ~ dnorm(0, 0.5),
    beta.food ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = fox)

precis(fox2)

plot(precis(fox2))
```

Now area has weakly positve and food has weakly negative effect (?) on weight.  not very reliable effects, however.

What if I add groupsize as well (I think this will muck things up)
```{r}
fox2b <- quap(
  alist(
    W ~ dnorm(mu, sigma),
    mu <- beta.area*A + beta.food*F + beta.group*G, 
    beta.area ~ dnorm(0, 0.5),
    beta.food ~ dnorm(0, 0.5),
    beta.group ~ dnorm(0, 0.5),
    sigma ~ dexp(1)
  ),
  data = fox)

precis(fox2b)

plot(precis(fox2b))

```
Hey, this looks perfectly reasonable.  now area and food positively associated with weight, and group size is strongly negatively associated.  Why did this work?  I have included things highly correlated with each other (group size and food); isn't this a problem?

```{r}

cor(fox$groupsize, fox$avgfood) #0.9014829
cor(fox$groupsize, fox$area) # 0.8275945
cor(fox$avgfood, fox$area) # 0.8831038
```




